Standard Candle Cosmology With Incomplete Spectroscopic Classification

### 2015-10-18 doc/article.tex
fix errors in likelihood

\documentclass[preprint,3p]{elsarticle}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}

@@

\subsection{Likelihood}

\subsubsection{Measurements}
In the following discussion,
spectroscopic and galaxy measurements are assumed
to provide prefect determinations of transient type and parameters, and
galaxy redshifts and parameters.  The working group is striving to get spectroscopic
redshifts of all potential hosts.  People who are interested in SN spectral features
or host-galaxy correlations (Chris, Rachel, Mat) can develop these models.

@@

\subsubsection{Constructing the Likelihood}

@@

The following is a sketch of the likelihood calculation for an example 
photometric selection criterion, which should serve as a useful guide once the criteria are fixed.
The criterion is for $\ge N$ photometric points
above a signal-to-noise threshold (translating to a distinct limiting signal $ADU_{0i}$
for measurement $i$ of a given transient) given $M$ available measurements. 
The fail/pass status of the measurements are stored as 0/1 in an array $b$.
There are $\sum_{k=N}^M \binom{M}{k}$ 
distinct $b$'s that allow acceptance of the light curve.

\subsubsection{Likelihood for the {\it obs} Set}
The likelihood for the {\it obs} set is
\begin{multline}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
= \sum_i \int_L dL\, p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | b_i, L, \ldots)
P(b_{i} |L, \ldots) P(L|\ldots) ,
\end{multline} 
where $L$ are the latent parameters.
The first term is the likelihood for the data given the condition for the  $b_i$ configuration
of threshold pass/fails  and all parameters.  The second term and
third terms are the probabilities of achieving the configuration
and latent parameters
respectively.

The likelihood for one of the $b$ configurations is derived from the likelihood without the configuration condition
\begin{equation}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | b_i, L, \ldots)
%=\frac{
% \epsilon(\mathit{ADU}_{SN}; b_i)
%  p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots)}
%  {\sum_{T_S} \int d\mathit{ADU}_{SN}\, dz_S\, d\theta_S\, d\theta_G\, 
% \epsilon(\mathit{ADU}_{SN}; b_i)
%  p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots)}\\
  =  \frac{ \prod_j
 \epsilon(\mathit{ADU}_{SN,j}; b_{ij})
  p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots)}
  {\prod_j P(b_{ij}{ |L,\dots)}} ,
\label{obsconf:eqn}
\end{equation}
where $j$ indexes photometric points and
$\epsilon$ is the efficiency of acceptance of that point from the underlying distribution.
For flux thresholds, the efficiency is
\begin{equation}
\epsilon(\mathit{ADU}_{SN,j}; b_{ij})= \mathcal{H}\left((-1)^{b_{ij}+1}(ADU_{SN,j}-ADU_{0,j})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{ij}=1$ the step in the efficiency
is up and for $b_{ij}=0$ the step is down.

The likelihood without the configuration condition is
\begin{multline}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots) \\
= \delta(z_S - g_z) \delta(\theta_S-\theta_T) \delta(\theta_G -g) \delta(T_S-T)  \mathcal{N}
\left(\mathit{ADU}_{SN} -\overline{\mathit{ADU}}_j(L ,\ldots), C_P \right),
\end{multline}
where $\mathcal{N}$ is an $M$-dimensional Normal distribution
and $\sigma$ is the photometric uncertainty.

@@

Assembling this together, the likelihood for the {\it obs} Set is
\begin{multline}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
= \sum_i \int_L dL\, 
\prod_j
 \epsilon(\mathit{ADU}_{SN,j}; b_{ij}) 
\mathcal{N}\left(\mathit{ADU}_{SN} - \overline{\mathit{ADU}}_j(L ,
g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots), C_P \right)\\
P(L|g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots),
\end{multline} 
where the latent parameters $T$, $\theta_T$ are not integrated over.

\subsubsection{Likelihood for the {\it mis} Set}
The likelihood for the {\it mis} set follow the same pattern as that of the
{\it obs} set except for not having to account for spectroscopic data.
\begin{multline}
p(\mathit{ADU}_{SN}, \theta_{G_1} |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g) \\
 = \sum_i \int_L dL\, 
\prod_j
 \epsilon(\mathit{ADU}_{SN,j}; b_{ij}) \sum_k 
\pi_k\mathcal{N}\left(\mathit{ADU}_{SN} - \overline{\mathit{ADU}}_j(L ,
g=\theta_{G_k}, \ldots), C_P \right)
P(L|g=\theta_{G_k}, \ldots), 
\end{multline} 
where $\pi_k$ is the probability that the $k$'th galaxy is the host, provided by Ravi.

>>>>

\documentclass[preprint,3p]{elsarticle}
%\documentclass[preprint]{aastex}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}

@@

\subsection{Measurements}
\label{measurements:sec}
In the following discussion,
spectroscopic and galaxy measurements are assumed
to provide prefect determinations of transient type and parameters, and
galaxy redshifts and parameters. 
\begin{align}
p(T_S|T) &= \delta(T_S-T),\\
p(\theta_S|\theta_T) &= \delta(\theta_S|\theta_T),\\
p(\theta_G|g) &= \delta(\theta_G-g)\\
p(z_S|g_z) &= \delta(z_S-g_z)\\.
\label{specz:eqn}
\end{align}
The working group is striving to get spectroscopic
redshifts of all potential hosts.  People who are interested in SN spectral features
or host-galaxy correlations (Chris, Rachel, Mat) can develop these models.

@@

\subsection{Constructing the Likelihood}

@@

The following is a sketch of the likelihood calculation for an example 
photometric selection criterion, which should serve as a useful guide once the criteria are fixed.
The criterion is for $\ge N$ photometric points
above a signal-to-noise threshold (translating to a distinct limiting signal $ADU_{0i}$
for measurement $i$ of a given transient) given $M$ available measurements. 
The fail/pass status of the measurements are stored as 0/1 in an array $b$.
There are $\sum_{k=N}^M \binom{M}{k}$ 
distinct $b$'s that allow acceptance of the light curve.  In the notation $b_{ijk}$, $i$ indexes
the configuration, $j$ the transient, and $k$ the photometric point of transient $j$.

\subsubsection{Likelihood for the {\it obs} Set}
The likelihood for the {\it obs} set is
\begin{multline}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
= \sum_i \int_L dL\, p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G , b_i|  L, \ldots)
 P(L|\ldots) \\
 \propto \sum_i  \epsilon(\mathit{ADU}_{SN}; b_{i}) \int_L dL\, p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G|  L, \ldots)P(L|\ldots).
\end{multline} 
It is proportional to the truncated likelihood of the underlying sample, where the truncation
is achieved through $\epsilon(\mathit{ADU}_{SN}; b_{i})$, the efficiency that a light curve is 
selected in for a given configuration.  For the limiting flux criteria
\begin{equation}
\epsilon(\mathit{ADU}_{SN}; b_{i})= \prod_{jk}\mathcal{H}\left((-1)^{b_{ijk}+1}(ADU_{SN,jk}-ADU_{0,jk})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{ijk}=1$ the step in the efficiency
is up and for $b_{ijk}=0$ the step is down.

%The likelihood without the configuration condition for our measurements is
%\begin{multline}
%p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots) \\
%= \delta(z_S - g_z) \delta(\theta_S-\theta_T) \delta(\theta_G -g) \delta(T_S-T)  \mathcal{N}
%\left(\mathit{ADU}_{SN} -\overline{\mathit{ADU}}_j(L ,\ldots), C_P \right).
%\end{multline}. 
%

@@

For measurements described in \S\ref{measurements:sec}, the likelihood for the {\it obs} Set is
\begin{multline}
p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
\propto \sum_i  \epsilon(\mathit{ADU}_{SN}; b_{i})  \int_L dL\, 
\mathcal{N}\left(\mathit{ADU}_{SN} - \overline{\mathit{ADU}}_j(L ,
g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots), C_P \right)\\
P(L|g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots),
\end{multline} 
where the latent parameters $T$, $\theta_T$ are not integrated over.
The normalization factor is parameter-dependent and is set by the condition
\begin{equation}
\sum_{T_S} \int d\mathit{ADU}_{SN}\,d{{z}}_S\,d\theta_{S}\,d\theta_G\,p(\mathit{ADU}_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \ldots)=1.
\end{equation}

\subsubsection{Likelihood for the {\it mis} Set}
The likelihood for the {\it mis} set follow the same pattern as that of the
{\it obs} set except for not having to account for spectroscopic data.
\begin{multline}
p(\mathit{ADU}_{SN}, \theta_{G_1} |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g) \\
 \propto \sum_i  \epsilon(\mathit{ADU}_{SN}; b_{i}) 
\int_L dL\, 
\sum_k 
\pi_k\mathcal{N}\left(\mathit{ADU}_{SN} - \overline{\mathit{ADU}}_j(L ,
g=\theta_{G_k}, \ldots), C_P \right)
P(L|g=\theta_{G_k}, \ldots), 
\end{multline} 
where $\pi_k$ is the probability that the $k$'th galaxy is the host, provided by Ravi.
